---
title: "P8105_hw1_tbd2117"
author: "Thiago de Araujo"
date: "9/13/2020"
output: github_document
---

### Problem 1

Here's the first code chunk to create the requested **dataframes** and its **means**:

```{r chunk1}
# random sample of size 10 from a standard Normal distribution
samp= rnorm(10)

# logical vector indicating whether elements of the sample are greater than 0
logic= c(samp > 0)

# character vector of length 10
char= c("This", "is", "a", "vector", "that", "contains", "a", "length", "of", "ten")

# factor vector of length 10, with 3 different factor “levels”
factor= factor(c("low", "medium", "high", "medium", "medium", "low", "medium", "low", "high", "low"))

# attempt to take the mean of each data variable
mean(samp)
mean(logic)
mean(char)
mean(factor)
````

Taking the mean of each variable works for the _samp_ variable (mean = `r mean(samp)`) and for the _logic_ variable (mean = `r mean(logic)`). It does not work for the char and factor variables.

```{r chunk1.5, eval=FALSE}
# appling the as.numeric function
as.numeric(logic)
as.numeric(char)
as.numeric(factor)
````
Using the _as.numeric_ function it is possible to see values being attributed to TRUE and FALSE in the _logic_ variable and for each level in the _factor_ variable. It does help explain what happens when trying to take means.

Here's the **second code chunk**:
```{r chunk2}
as.numeric(logic) * samp
as.factor(logic) * samp
as.numeric(as.factor(logic)) * samp
```

### Problem 2

```{r}
# loading the penguins dataset
data("penguins", package = "palmerpenguins")
y= as.numeric(penguins$flipper_length_mm)
````

The _penguins_ dataset contains data on penguin species (`r levels(penguins$species)`), islands (`r levels(penguins$island)`), length of bills and flippers, body mass, sex, and year. 
The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
The mean flipper length is `r mean(y) `.

```{r message=FALSE}
# calling library
library(tidyverse)

# scatterplot of flipper vs. bill length
ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species)) + geom_point()

# exporting scatterplot 
ggsave("scatter_plot.pdf", height = 4, width = 6)
```